---
title: "How to use `rvdat`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{How to use `rvdat`}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette is a working draft. There will be lots of changes as `rvdat` gets fleshed out, so please check back often!

The most recent changes were made on `r Sys.Date()`.

## Getting started

```{r setup}
library(rvdat)
```

First, as we should always do when beginning a `rvdat` workflow, we tell R where our `vdat.exe` is located. Mine is within an installed version of [Fathom Connect](https://support.fishtracking.innovasea.com/s/downloads).

```{r, vdat-here}
vdat_here("c:/program files/innovasea/fathom connect/vdat.exe")
```

If you don't want to do this every time you re-open R, you can add `VDAT_EXE="PATH_TO_VDAT.EXE"` to your `.Renviron` file. `rvdat` will automatically look for it there when you restart R. You can do this manually, but I find that the easiest way to edit `.Renviron` is using [`usethis::edit_r_environ()`](https://usethis.r-lib.org/reference/edit.html).

Next, let's download a sample VRL file from the `glatos` package using `utils::download.file`. I'll be keeping everything in a temporary directory just to keep tidy.

```{r, dl-vrl}
td <- file.path(
  tempdir(),
  "vignette"
)

dir.create(
  td
)

download.file(
  url = file.path(
    "https://github.com/ocean-tracking-network/glatos/raw/dev/inst/extdata",
    "detection_files_raw/VR2AR_546310_20190613_1.vrl"
  ),
  destfile = file.path(
    td,
    "VR2AR_546310_20190613_1.vrl"
  ),
  mode = "wb",
  quiet = TRUE
)
```

## Check the file's metadata

To take a quick glimpse of the VDAT file's metadata, we can use `vdat_inspect`. Note that this only works for VDAT files (`.vrl` and `.vdat`), not any exported CSVs. `vdat_insepct` converts this information to a data.frame if you'd like to keep it around.

```{r}
inspect_df <- vdat_inspect(
  file.path(td, "VR2AR_546310_20190613_1.vrl")
)

head(inspect_df)
```

## Converting VDAT file to a CSV

Now that we have a VRL to play with and know something about it, let's convert it to a csv. It's pretty quick with `vdat_to_csv`.

```{r, vrl-to-csv}
vdat_to_csv(
  file.path(td, "VR2AR_546310_20190613_1.vrl"),
  outdir = td
)

list.files(td, pattern = "VR2AR")
```
`vdat` is very respectful of your data in that it never overwrites files that exist in the same directory. It'll just tack on a time stamp to the name of the newly-exported file.

```{r, vrl-to-csv2}
vdat_to_csv(
  file.path(td, "VR2AR_546310_20190613_1.vrl"),
  outdir = td
)

list.files(td, pattern = "VR2AR")
```
## Converting a VDAT file to a CSV with time correction

Adding the `time_corrected = TRUE` argument to `vdat_to_csv` corrects for clock drift using `vdat`'s internal algorithm:

```{r, vrl2csv-time-correct}
vdat_to_csv(
  file.path(td, "VR2AR_546310_20190613_1.vrl"),
  outdir = td,
  time_corrected = TRUE
)
```

A new column is added with time corrected down to the microsecond.

```{r}
corrected <- read.csv(
  list.files(td, pattern = "VR2AR.*\\]\\.csv$", full.names = TRUE)[2],
  header = FALSE
)

corrected[46, c(2:3)]
```


## Splitting a VDAT file into a folder of CSVs

As outlined in `vignette('vdat-data-structure')`, the flat CSV exported via `vdat_to_csv` is a mush-together of all available data types in the VDAT file. `vdat.exe` can split these into a folder of files according to data type for you, accessible via `rvdat` in a few ways:

  -   `vdat_to_folder("PATH_TO_FILE")`, which is a convenience wrapper around...
  -   `vdat_to_csv("PATH_TO_FILE", folder = TRUE)`, which is a convenience wrapper around...
  -   `vdat_call(c('convert', '--format=csv.fathom.split', 'PATH_TO_FILE'))`.

```{r}
vdat_to_folder(
  file.path(td, "VR2AR_546310_20190613_1.vrl"),
  outdir = td
)

list.dirs(td, full.names = F, recursive = F)

list.files(
  list.dirs(td, full.names = T, recursive = F)
)
```


## Other file types

Though we just used a VR2AR's `VRL` file here, this works for multiple different receiver styles and file types. For example, a HR-180 `VDAT` file:

```{r}
download.file(
  url = file.path(
    "https://github.com/ocean-tracking-network/glatos/raw/dev/inst/extdata",
    "detection_files_raw/HR2-180_461396_2021-04-20_173145.vdat"
  ),
  destfile = file.path(
    td,
    "HR2-180_461396_2021-04-20_173145.vdat"
  ),
  mode = "wb",
  quiet = TRUE
)

hr_output <- vdat_inspect(
  file.path(td, "HR2-180_461396_2021-04-20_173145.vdat")
)
head(hr_output)
```



```{r, clean-up, echo=FALSE}
unlink(td, recursive = TRUE)
```

