---
title: "How to use `rvdat`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{How to use `rvdat`}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Getting started

```{r setup}
library(rvdat)
```

First, as we should always do when beginning a `rvdat` workflow, we tell R where our `vdat.exe` is located. Mine is within an installed version of [Fathom Connect](https://support.fishtracking.innovasea.com/s/downloads).

```{r, vdat-here}
vdat_here("c:/program files/innovasea/fathom/vdat.exe")
```

Next, let's download a sample VRL file from the `glatos` package using `utils::download.file`. I'll be keeping everything in a temporary directory just to keep tidy.

```{r, dl-vrl}
td <- file.path(
  tempdir(),
  "vignette"
)

dir.create(
  td
)

download.file(
  url = file.path(
    "https://github.com/ocean-tracking-network/glatos/raw/dev/inst/extdata",
    "detection_files_raw/VR2AR_546310_20190613_1.vrl"
  ),
  destfile = file.path(
    td,
    "VR2AR_546310_20190613_1.vrl"
  ),
  mode = "wb",
  quiet = TRUE
)
```

## Check the file's metadata

To take a quick glimpse of the VDAT file's metadata, we can use `vdat_inspect`. Note that this only works for VDAT files (`.vrl` and family), not any exported CSVs.

```{r}
vdat_inspect(
  file.path(td, "VR2AR_546310_20190613_1.vrl")
)
```

## Converting a VRL to a CSV

Now that we have a VRL to play with and know something about it, let's convert it to a csv. It's pretty quick with `vdat_to_csv`.

```{r, vrl-to-csv}
vdat_to_csv(
  file.path(td, "VR2AR_546310_20190613_1.vrl"),
  outdir = td
)

list.files(td, pattern = "VR2AR")
```
`vdat` is very respectful of your data in that it never overwrites files that exist in the same directory. It'll just tack on a time stamp to the name of the newly-exported file.

```{r, vrl-to-csv2}
vdat_to_csv(
  file.path(td, "VR2AR_546310_20190613_1.vrl"),
  outdir = td
)

list.files(td, pattern = "VR2AR")
```
## Converting a VRL to a CSV with time correction

Adding the `time_corrected = TRUE` argument to `vdat_to_csv` corrects for clock drift using `vdat`'s internal algorithm:

```{r, vrl2csv-time-correct}
vdat_to_csv(
  file.path(td, "VR2AR_546310_20190613_1.vrl"),
  outdir = td,
  time_corrected = TRUE
)
```



## Converting a VRL to a CSV folder

As outlined in `vignette('vdat-data-structure')`, the flat CSV exported via `vdat_to_csv` is a mush-together of all available data types in the VDAT file. `vdat.exe` can split these into a folder of files according to data type for you, accessible using `rvdat` via `vdat_to_folder` or `vdat_call(c('convert', '--format=csv.fathom.split', 'FILENAME'))`.

```{r}
vdat_to_folder(
  file.path(td, "VR2AR_546310_20190613_1.vrl"),
  outdir = td
)

list.dirs(td, full.names = F, recursive = F)

list.files(
  list.dirs(td, full.names = T, recursive = F)
)
```





```{r, clean-up, echo=FALSE}
unlink(td, recursive = TRUE)
```

